### 2023-01-14 draft.tex
Deploying to gh-pages from @ gottacatchenall/ms_false_negatives@c62b1bbd14631093c54d29eb7c03d729d3efe513 🚀

%
\href{https://orcid.org/0000-0002-6506-6487}{Michael D.\,Catchen}%
%
\,\textsuperscript{1,2}\quad %
\href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
%
\,\textsuperscript{3,2}\quad %
\href{https://orcid.org/0000-0002-6004-4027}{Laura\,Pollock}%
%
\,\textsuperscript{1,2}\quad %
\href{https://orcid.org/0000-0001-6075-8081}{Andrew\,Gonzalez}%
%
\,\textsuperscript{1,4}

@@

        {\bfseries Abstract:}\,Ecosystems are composed of networks of
interacting species. These interactions allow communities of species to
persist through time through both neutral and adaptive processes. Still
a robust understanding of (and ability to predict and forecast)
interactions among species remains elusive. This knowledge-gap is
largely driven by a shortfall of data---although species occurrence data
has rapidly increased in the last decade, species interaction data has
not kept pace, largely due to the intrinsic difficulty and effort
required to sample interactions. These sampling challenges bias data and
hinder inferences about the structure and dynamics of interactions
networks. Here, we demonstrate the realized false-negative rate (the
percentage of species that actually interact but for which we do not yet
have a record) can be quite high, even in thoroughly sampled systems,
due to the intrinsic variation in abundances across species in a
community. We illustrate how a null model of occurrence detection can be
used to estimate the false-negative rate in a given dataset. We also
show how to directly incorporate uncertainty due to observation error
into model-based predictions of interaction probabilities between
species. One hypothesis is that interactions between ``rare'' species
are themselves rare because these species are less likely to encounter
one-another than species of higher relative abundance. However, we
demonstrate that across several datasets of spatial or temporally
replicated networks, there are positive associations that suggest these
interactions actually exist but just are not observed. Finally, we
assess how false negatives influence various models of network
prediction, and recommend directly accounting for observation error in
predictive models. We conclude by discussing how the understanding of
false-negatives can inform how we design monitoring schemes for species
interactions.\\%

@@

Species interactions drive many processes in evolution and ecology. A
better understanding of species interactions is an imperative to
understand the evolution of life on Earth, to mitigate the impacts of
anthropogenic change on biodiversity (Makiola \emph{et al.} 2020), and
for predicting zoonotic spillover of disease to prevent future pandemics
(Becker \emph{et al.} 2021). At the moment we lack sufficient data to
meet these challenges (Poisot \emph{et al.} 2021), largely because
species interactions are hard to sample (Jordano 2016). Over the past
few decades biodiversity data has become increasingly available through
remotely collected data and adoption of open data practices (Kenall
\emph{et al.} 2014; Stephenson 2020). Still, interaction data remains
relatively scarce because sampling typically requires human observation.
This induces a constraint on the amount, spatial scale, and temporal
frequency of resulting data that it is feasible to collect by humans.
Many crowdsourced methods for biodiversity data aggregation (e.g.~GBIF,
eBird) still relies on automated identification of species, which does
not easily generalize to interaction sampling. There is interest in
using remote methods for interaction sampling, which primarily detect
co-occurrence and derive properties like species avoidance from this
data (Niedballa \emph{et al.} 2019). However, this itself is not
necessarily indicative of an interaction (Blanchet \emph{et al.} 2020).
This is an example of semantic confusion around the word
``interaction''---for example one might consider competition a type of
species interaction, even though it is marked by a lack of co-occurrence
between species, unlike other types of interactions, like trophism or
pollination, which require both species to be together at the same place
and time. Here we consider interaction in the latter sense, where two
species have fitness consequences on one-another if (and only if) they
are in the sample place at the same time. In addition, here we only
consider direct (not higher-order) interactions.

We cannot feasibly observe all (or even most) of the interactions that
occur in an ecosystem. This means we can be confident two species
actually interact if we have a record of it (given an estimate of
species misidentification probability), but not at all confident that a
pair of species \emph{do not} interact if we have \emph{no record} of
those species observed together. In other words, it is difficult to
distinguish \emph{true-negatives} (two species never interact) from
\emph{false-negatives} (two species interact sometimes, but we do not
have a record of it). For a concrete example of a false-negative in a
food web, see fig.~\ref{fig:concept}. Because even the most highly
sampled systems will still contain missing interactions, there is
increasing interest in combining species-level data (e.g.~traits,
abundance, range, phylogenetic relatedness, etc.) to build models to
predict interactions between species we haven't observed together before
(Strydom \emph{et al.} 2021). However, the noise of false-negatives
could impact the efficacy of our predictive models and have practical
consequences for answering questions about interactions (de Aguiar
\emph{et al.} 2019). This data constraint is amplified as the
interaction data we have is geographically biased toward the usual
suspects (Poisot \emph{et al.} 2021). We therefore need a statistical
approach to assessing these biases in the observation process and their
consequences for our understanding of interaction networks.

The importance of \emph{sampling effort} and its impact on resulting
ecological data has produced a rich body of literature.The recorded
number of species in a dataset or sample depends on the total number of
observations (Walther \emph{et al.} 1995; Willott 2001), as do estimates
of population abundance (Griffiths 1998). This relationship between
sampling effort and spatial coverage and species detectability . This
has motivated more quantitatively robust approaches to account for error
in sampling data in many contexts: to determine if a given species is
extinct (Boakes \emph{et al.} 2015), to determine sampling design (Moore
\& McCarthy 2016), and to measure species richness across large scales
(Carlson \emph{et al.} 2020). In the context of interactions, an initial
concern was the compounding effects of limited sampling effort combined
with the amalgamation of data (across both study sites, time of year,
and taxonomic scales) could lead any empirical set of observations to
inadequately reflect the reality of how species interact (Paine 1988) or
the structure of the network as a whole (Martinez \emph{et al.} 1999;
McLeod \emph{et al.} 2021). Martinez \emph{et al.} (1999) showed that in
a plant-endophyte trophic network, network connectance is robust to
sampling effort, but this was done in the context of a system for which
observation of 62,000 total interactions derived from 164,000
plant-stems was feasible. In some systems (e.g.~megafauna food-webs)
this many observations is either impractical or infeasible due to the
absolute abundance of the species in question.

The intrinsic properties of ecological communities create several
challenges for sampling: first, species are not observed with equal
probability---we are much more likely to observe a species of high
abundance than one of very low abundance (Poisot \emph{et al.} 2015).
Canard \emph{et al.} (2012) presents a null model of food-web structure
where species encounter one-another in proportion to each species'
relative-abundance. This assumes that there are no associations in
species co-occurrence due to an interaction (perhaps because this
interaction is ``important'' for both species; Cazelles \emph{et al.}
(2016)), but in this paper we later show increasing strength of
associations leads to increasing probability of false-negatives in
interaction data, and that these positive associations are rampant in
existing network data. Second, observed co-occurrence is often equated
with meaningful interaction strength, but this is not necessarily the
case (Blanchet \emph{et al.} 2020)---a true ``non-interaction'' would
require that neither of two species, regardless of whether they
co-occur, ever exhibit any meaningful effect on the fitness of the
other. So, although co-occurrence is not directly indicative of an
interaction, it \emph{is} a precondition for an interaction.

Here, we illustrate how our confidence that a pair of species never
interacts highly depends on sampling effort. We suggest that surveys of
species interactions can benefit from simulation modeling of the
sampling process. We demonstrate how the realized false-negative rate of
interactions is related to the relative abundance of the species pool,
and use simulation to produce a null estimate of the false-negative rate
given total sampling effort (the total count of all individuals of all
species seen), and to introduce a method for introducing uncertainty
into model predictions of interaction probability to account for
observation error. We then show that positive associations in
co-occurrence data can increase the realized number of false-negatives,
and demonstrate these positive associations are rampant in network
datasets, and conclude by recommending that the simulation of sampling
effort and species occurrence can and should be used to help design
surveys of species interaction diversity (Moore \& McCarthy 2016), and
by advocating use of null models like those presented here as a tool for
both guiding design of surveys of species interactions and for modeling
detection error in predictive models.

\begin{figure}
\hypertarget{fig:concept}{%
\centering
\includegraphics{./figures/concept.png}
\caption{This conceptual example considers a sample of the trophic
community of bears, wolves, salmon (pink fish), pike (yellow fish),
berry trees, and aspen trees. The true metaweb (all realized
interactions across the entire spatial extent) is shown on the left. In
the center is what a hypothetical ecologist samples at each site. Notice
that although bears are observed co-occurring with both salmon and pike,
there was never a direct observation of bears eating pike, even though
they actually do. Therefore, this interaction between bears and pike is
a false negative.}\label{fig:concept}
}
\end{figure}

@@

@@

We start with a naive model of interaction detection: we assume that
every interacting pair of species is incorrectly observed as
not-interacting with an independent and fixed probability, which we
denote \(p_{fn}\) and subsequently refer to as the False-Negative Rate
(FNR). If we observe the same species not-interacting \(N\) times, then
the probability of a true-negative (denoted \(p_{tn}\)) is given by
\(p_{tn}=1-(p_{fn})^N\). This relation (the probability-mass-function of
geometric distribution, a special case of the negative-binomial
distribution) is shown in fig.~\ref{fig:geometric}(A) for varying values
of \(p_{fn}\) and illustrates a fundamental link between our ability to
reliably say an interaction doesn't exist---\(p_{tn}\)---and the number
of times \(N\) we have observed a given species. In addition, note that
there is no non-zero \(p_{fn}\) for which we can ever \emph{prove} that
an interaction does not exist---no matter how many observations of
non-interactions \(N\) we have, \(p_{tn}<1\).

@@

Because the true FNR is latent, we can never actually be sure what the
actual number of false negatives in our data---however, we can use
simulation to estimate it for datasets of a given size using neutral
models of observation. If some of the ``worst-case'' FNRs presented in
fig.~\ref{fig:geometric}(A) seem unrealistically high, consider that
species are observed in proportion to their relative abundance. In the
next section we demonstrate that the distribution of abundance in
ecosystems can lead to very high realized values of FNR (\(p_{fn}\))
simply as an artifact of sampling effort.

\begin{figure}
\hypertarget{fig:geometric}{%
\centering
\includegraphics{./figures/fig1.png}
\caption{\textbf{(A)} The probability that an observed interaction is a
true negative (y-axis) given how many times it has been sampled as a
non-interaction (x-axis). Each color reflects a different value of
\(p_{fn}\), the false-negative rate (FNR)---this is effectively the cdf
of the geometric distribution. \textbf{(B)} The expected number of total
observations needed (colors) to observe 10 co-occurrences between a
species with relative abundance \(P(A)\) (x-axis) and a second species
with relative abundance \(P(Y)\). \textbf{(C)}: False negative rate
(y-axis) as a function of total sampling effort (x-axis) and network
size, computed using the method described above. For 500 independent
draws from the niche model (Williams \& Martinez (2000)) at varying
levels of species richness (colors) with connectance drawn according to
the flexible-links model (MacDonald \emph{et al.} (2020)) as described
in the main text. For each draw from the niche model, 200 sets of 1500
observations are simulated, for which each the mean false negative rate
at each observation-step is computed. Means denoted with points, with 1
in the first shade and 2 in the second. \textbf{(D)}: Same as
\textbf{(C)}, except using empirical food webs from Mangal database,
where richness. The outlier on \textbf{(D)} is a 714 species
food-web.}\label{fig:geometric}
}
\end{figure}

@@

where the MAP estimate of \((\mu, \phi)\) applied to Mangal data from
(MacDonald \emph{et al.} 2020) is \((\mu=0.086, \phi=24.3)\). All
simulations were done with 500 independent replicates of unique niche
model networks per unique number of observations \(n\). All analyses
presented here are done in Julia v1.8 (Bezanson \emph{et al.} 2015)
using both EcologicalNetworks.jl v0.5 and Mangal.jl v0.4 (Banville
\emph{et al.} 2021) and are hosted on
\href{https://github.com/gottacatchenall/ms_false_negatives/tree/main/src}{Github}).
Note that the empirical data, for the reasons described above, very
likely already contains many false negatives, we'll revisit this issue
in the final section.

From fig.~\ref{fig:geometric}(C) it is evident that the number of
species considered in a study is inseparable from the false-negative
rate in that study, and this effect should be taken into account when
designing samples of ecological networks in the future. We see a similar
qualitative pattern in fig.~\ref{fig:geometric}(D) where the FNR drops
off quickly as a function of observation effort, mediated by total
richness. The practical consequence of the bottom row of
fig.~\ref{fig:geometric} is whether the total number of observations of
all species (the x-axis) for the threshold FNR we deem acceptable (the
y-axis) is feasible. This raises two points: first, empirical data on
interactions are subject to the practical limitations of funding and
human-work hours, and therefore existing data tend to fall on the order
of hundreds or thousands observations of individuals per site. Clear
aggregation of data on sampling effort has proven difficult to find and
a meta-analysis of network data and sampling effort seems both pertinent
and necessary, in addition to the effects of aggregation of interactions
across taxonomic scales (Gauzens \emph{et al.} 2013; Giacomuzzo \&
Jordán 2021). This inherent limitation on in-situ sampling means we
should optimize where we sample across space so that for a given number
of samples, we obtain the maximum information possible. Second, what is
meant by ``acceptable'' FNR? This raises the question: does a shifting
FNR lead to rapid transitions in our ability inference and predictions
about the structure and dynamics of networks, or does it produce a
roughly linear decay in model efficacy? We explore this in the next
section.

@@

\begin{figure}
\hypertarget{fig:resampling_concept}{%
\centering
\includegraphics{./figures/uncertainty_sampler.png}
\caption{(A) The process for estimating the false-negative rate (FNR)
for an interaction dataset consisting of \(N\) total observed
interactions. (B) The method for resampling interaction probability
based on estimates of false-negative and false-positive rates. (C) The
method for interaction probability resampling applied to three mammals
and three parasites from the Hadfield \emph{et al.} (2014)
dataset.}\label{fig:resampling_concept}
}
\end{figure}

@@

\hypertarget{positive-associations-in-co-occurrence-increase-the-false-negative-rate}{%
\section{Positive associations in co-occurrence increase the
false-negative
rate}\label{positive-associations-in-co-occurrence-increase-the-false-negative-rate}}

The model above doesn't consider the possibility that there are positive
or negative associations which shift the probability of species
cooccurrence together due to their interaction (Cazelles \emph{et al.}
2016). However, here we demonstrate that the probability of observing a
false-negative can be higher if there is some positive association in
the occurrence of species \(A\) and \(B\). If we denote the probability
that we observe the co-occurrence of two species \(A\) and \(B\) as
\(P(AB)\) and if there is no association between the marginal
probabilities of observing \(A\) and observing \(B\), denoted \(P(A)\)
and \(P(B)\) respectively, then the probability of observing their
co-occurrence is the product of the marginal probabilities for each
species, \(P(AB) = P(A)P(B)\). In the other case where there is some
positive strength of association between observing both \(A\) and \(B\)
because this interaction is ``important'' for each species, then the
probability of observation both \(A\) and \(B\), \(P(AB)\), is greater
than \(P(A)P(B)\) as \(P(A)\) and \(P(B)\) are not independent and
instead are positively correlated, i.e.~\(P(AB)> P(A)P(B)\). In this
case, the probability of observing a single false-negative in our naive
model from fig.~\ref{fig:geometric}(A) is \(p_{fn}= 1-P(AB)\), which due
to the above inequality implies \(p_{fn}>1-P(A)P(B)\). This indicates an
increasingly greater probability of a false negative as the strength of
association gets stronger, \(P(AB) \to P(AB) \gg P(A)P(B)\). However,
this still does not consider variation in species abundance in space and
time (Poisot \emph{et al.} 2015). If positive or negative associations
between species structure variation in the distribution of \(P(AB)\)
across space/time, then the spatial/temporal biases induced by data
collection would further impact the realized false negative rate, as the
probability of false negative would not be constant for each pair of
species across sites.

To test for these positive associations in data we scoured Mangal for
datasets with many spatial or temporal replicates of the same system.
For each dataset, we compute the marginal probability \(P(A)\) of
occurrence of each species \(A\) across all networks in the dataset. For
each pair of interacting species \(A\) and \(B\), we then compute and
compare the probability of co-occurrence if each species occurs
independently, \(P(A)P(B)\), to the empirical joint probability of
co-occurrence, \(P(AB)\). Following our analysis above, if \(P(AB)\) is
greater than \(P(A)P(B)\), then we expect our neutral estimates of the
FNR above to underestimate the realized FNR. In fig.~\ref{fig:mangal},
we see the difference between \(P(AB)\) and \(P(A)P(B)\) for the seven
suitable datasets with enough spatio-temporal replicates and a shared
taxonomic backbone (meaning all individual networks use common species
identifiers) found on Mangal to perform this analysis. Further details
about each dataset are reported in tbl.~\ref{tbl:id}.

@@

Here, we assess the effect of false negatives on our ability to make
predictions about interactions, as well as their effect on network
structure. The prevalence of false-negatives in data is the catalyst for
interaction prediction in the first place, and as a result methods have
been proposed to counteract this bias (Stock \emph{et al.} 2017; Poisot
\emph{et al.} 2022). However, it is feasible that the FNR in a given
dataset is so high that it could induce too much noise for an
interaction prediction model to detect the signal of possible
interaction between species.

@@

\begin{figure}
\hypertarget{fig:addedfnr}{%
\centering
\includegraphics{./figures/fig3.png}
\caption{\textbf{(A)} The area-under the receiver-operator curve
(ROC-AUC) and \textbf{(B)} The area-under the precision-recall curve
(PR-AUC; right) for each different predictive model (colors/shapes)
across a spectrum of the proportion of added false negatives (x-axis).
\textbf{(C)} The mean trophic-level of all species in a network
generated with the niche model across different species richnesses
(colors). For each value of the FNR, the mean trophic level was computed
across 50 replicates. The shaded region for each line is one
standard-deviation across those replicates.}\label{fig:addedfnr}
}
\end{figure}

>>>>

%
\href{https://orcid.org/0000-0002-6506-6487}{Michael D.\,Catchen}%
%
\,\textsuperscript{1,2}\quad %
\href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
%
\,\textsuperscript{3,2}\quad %
\href{https://orcid.org/0000-0002-6004-4027}{Laura J.\,Pollock}%
%
\,\textsuperscript{1,2}\quad %
\href{https://orcid.org/0000-0001-6075-8081}{Andrew\,Gonzalez}%
%
\,\textsuperscript{1,4}

@@

        {\bfseries Abstract:}\,Ecosystems are composed of networks of
interacting species. These interactions allow communities of species to
persist through time through both neutral and adaptive processes.
Despite their importance, a robust understanding of (and ability to
predict and forecast) interactions among species remains elusive. This
knowledge-gap is largely driven by a shortfall of data---although
species occurrence data has rapidly increased in the last decade,
species interaction data has not kept pace, largely due to the intrinsic
difficulty and effort required to sample interactions. This means there
are many interactions between species that occur in nature, but we do
not think this interaction occurs because we have no record of it. These
so-called ``false-negatives'' bias data and hinder inference about the
structure and dynamics of interaction networks. Here, we demonstrate the
realized rate of false-negatives in data can be quite high, even in
thoroughly sampled systems, due to the intrinsic variation in abundances
across species in a community. We illustrate how a null model of
occurrence detection can be used to estimate the false-negative rate in
a given dataset. We also show how to directly incorporate uncertainty
due to observation error into model-based predictions of interaction
probabilities between species. One hypothesis is that interactions
between ``rare'' species are themselves rare because these species are
less likely to encounter one-another than species of higher relative
abundance, and that this can (in part) explain the common pattern of
nestedness in bipartite interaction networks. However, we demonstrate
that across several datasets of spatial or temporally replicated
networks, there are positive associations between species co-occurrence
and interactions, which suggests these interactions among ``rare''
species actually exist but simply are not observed. Finally, we assess
how false negatives influence various models of network prediction, and
recommend directly accounting for observation error in predictive
models. We conclude by discussing how the understanding of
false-negatives can inform how we design monitoring schemes for species
interactions.\\%

@@

Species interactions drive many processes in evolution and ecology. A
better understanding of species interactions is an imperative to
understand the evolution of life on Earth, to mitigate the impacts of
anthropogenic change on biodiversity (Makiola \emph{et al.} 2020), and
for predicting zoonotic spillover of disease to prevent future pandemics
(Becker \emph{et al.} 2021). At the moment we lack sufficient data to
meet these challenges (Poisot \emph{et al.} 2021), largely because
species interactions are hard to sample (Jordano 2016). Over the past
few decades biodiversity data has become increasingly available through
remotely collected data and adoption of open data practices (Kenall
\emph{et al.} 2014; Stephenson 2020). Still, interaction data remains
relatively scarce because sampling typically requires human observation.
This induces a constraint on the amount, spatial scale, and temporal
frequency of resulting data that it is feasible to collect by humans.
Many crowdsourced methods for biodiversity data aggregation (e.g.~GBIF,
eBird) still rely on automated identification of species, which does not
easily generalize to interaction sampling. There is interest in using
remote methods for interaction sampling, which primarily detect
co-occurrence and derive properties like species avoidance from this
data (Niedballa \emph{et al.} 2019). However, co-occurrence itself is
not necessarily indicative of an interaction (Blanchet \emph{et al.}
2020). This is an example of semantic confusion around the word
``interaction''---for example one might consider competition a type of
species interaction, even though it is marked by a lack of co-occurrence
between species, unlike other types of interactions, like predation or
parasitism, which require both species to be together at the same place
and time. Here we consider interaction in the latter sense, where two
species have fitness consequences on one-another if (and only if) they
are in the sample place at the same time. In addition, here we only
consider direct (not higher-order) interactions.

We cannot feasibly observe all (or even most) of the interactions that
occur in an ecosystem. This means we can be confident two species
actually interact if we have a record of it (assuming they are correctly
identified), but not at all confident that a pair of species \emph{do
not} interact if we have \emph{no record} of those species observed
together. In other words, it is difficult to distinguish
\emph{true-negatives} (two species never interact) from
\emph{false-negatives} (two species interact sometimes, but we do not
have a record of this interaction). For a concrete example of a
false-negative in a food web, see fig.~\ref{fig:concept}. Because even
the most highly sampled systems will still contain false-negatives,
there is increasing interest in combining species-level data
(e.g.~traits, abundance, range, phylogenetic relatedness, etc.) to build
models to predict interactions between species we haven't observed
together before (Strydom \emph{et al.} 2021). However, the noise of
false-negatives could impact the efficacy of our predictive models and
have practical consequences for answering questions about interactions
(de Aguiar \emph{et al.} 2019). This data constraint is amplified as the
interaction data we have is geographically biased toward the usual
suspects (Poisot \emph{et al.} 2021). We therefore need a statistical
approach to assessing these biases in the observation process and their
consequences for our understanding of interaction networks.

The importance of \emph{sampling effort} and its impact on resulting
ecological data has produced a rich body of literature.The recorded
number of species in a dataset or sample depends on the total number of
observations (Walther \emph{et al.} 1995; Willott 2001), as do estimates
of population abundance (Griffiths 1998). This relationship between
sampling effort and spatial coverage and species detectability has
motivated more quantitatively robust approaches to account for error in
sampling data in many contexts: to determine if a given species is
extinct (Boakes \emph{et al.} 2015), to determine sampling design (Moore
\& McCarthy 2016), and to measure species richness across large scales
(Carlson \emph{et al.} 2020). In the context of interactions, an initial
concern was the compounding effects of limited sampling effort combined
with the amalgamation of data (across both study sites, time of year,
and taxonomic scales) could lead any empirical set of observations to
inadequately reflect the reality of how species interact (Paine 1988) or
the structure of the network as a whole (Martinez \emph{et al.} 1999;
McLeod \emph{et al.} 2021). Martinez \emph{et al.} (1999) showed that in
a plant-endophyte trophic network, network connectance is robust to
sampling effort, but this was done in the context of a system for which
observation of 62,000 total interactions derived from 164,000
plant-stems was feasible. In some systems (e.g.~megafauna food-webs)
this many observations is either impractical or infeasible due to the
absolute abundance of the species in question.

The intrinsic properties of ecological communities create several
challenges for sampling: first, species are not observed with equal
probability---we are much more likely to observe a species of high
abundance than one of very low abundance (Poisot \emph{et al.} 2015).
Canard \emph{et al.} (2012) presents a null model of food-web structure
where species encounter one-another in proportion to each species'
relative-abundance. This assumes that there are no associations in
species co-occurrence due to an interaction (perhaps because this
interaction is ``important'' for both species; Cazelles \emph{et al.}
(2016)), but in this paper we later show increasing strength of
associations leads to increasing probability of false-negatives in
interaction data, and that these positive associations are common in
existing network data. Second, observed co-occurrence is often equated
with meaningful interaction strength, but this is not necessarily the
case (Blanchet \emph{et al.} 2020)---a true ``non-interaction'' would
require that neither of two species, regardless of whether they
co-occur, ever exhibit any meaningful effect on the fitness of the
other. So, although co-occurrence is not directly indicative of an
interaction, it \emph{is} a precondition for an interaction.

Here, we illustrate how our confidence that a pair of species never
interacts highly depends on sampling effort. We suggest that surveys of
species interactions can benefit from simulation modeling of the
sampling process. We demonstrate how the realized false-negative-rate of
interactions is related to the relative abundance of the species pool,
and introduce a method to produce a null estimate of the
false-negative-rate given total sampling effort (the total count of all
interactions seen among all species-pairs) and a method for including
uncertainty into model predictions of interaction probabilities to
account for observation error. We then show that positive associations
in co-occurrence data can increase the realized number of
false-negatives, and demonstrate these positive associations are rampant
in network datasets, and conclude by recommending that the simulation of
sampling effort and species occurrence can and should be used to help
design surveys of species interaction diversity (Moore \& McCarthy
2016), and by advocating use of null models like those presented here as
a tool for both guiding design of surveys of species interactions and
for modeling detection error in predictive models.

\begin{figure}
\hypertarget{fig:concept}{%
\centering
\includegraphics{./figures/concept.png}
\caption{This conceptual example considers a sample of the trophic
community of bears, wolves, salmon (pink fish), pike (yellow fish),
berry trees, and aspen trees. The true metaweb (all realized
interactions across the entire spatial extent) is shown on the left. In
the center is what a hypothetical ecologist samples at each site. Notice
that although bears are observed co-occurring with both salmon and pike,
there was never a direct observation of bears eating pike, even though
they actually do. Therefore, this interaction between bears and pike is
a false-negative.}\label{fig:concept}
}
\end{figure}

@@

In this section, we demonstate how difference in relative-abundance can
lead to many false-negatives in interaction data. We also introduce a
method for producing a null estimate of the false-negative-rate in
datasets via simulation, and a method for incorporating uncertainty
directly into predictions of species interactions to account for
observation error.

@@

We start with a naive model of interaction detection: we assume that
every interacting pair of species is incorrectly observed as
not-interacting with an independent and fixed probability, which we
denote \(p_{fn}\) and subsequently refer to as the False-Negative-Rate
(FNR). If we observe the same species not-interacting \(N\) times, then
the probability of a true-negative (denoted \(p_{tn}\)) is given by
\(p_{tn}=1-(p_{fn})^N\). This relation (the probability-mass-function of
geometric distribution, a special case of the negative-binomial
distribution) is shown in fig.~\ref{fig:geometric}(A) for varying values
of \(p_{fn}\) and illustrates a fundamental link between our ability to
reliably say an interaction doesn't exist---\(p_{tn}\)---and the number
of times \(N\) we have observed a given species. In addition, note that
there is no non-zero \(p_{fn}\) for which we can ever \emph{prove} that
an interaction does not exist---no matter how many observations of
non-interactions \(N\) we have, \(p_{tn}<1\).

@@

Because the true FNR is latent, we can never actually be sure what the
actual number of false-negatives in our data---however, we can use
simulation to estimate the FNR for datasets of a given size using
neutral models of observation. If some of the ``worst-case'' FNRs
presented in fig.~\ref{fig:geometric}(A) seem unrealistically high,
considering that species are observed in proportion to their relative
abundance. In the next section we demonstrate that the distribution of
abundance in ecosystems can lead to very high realized values of FNR
(\(p_{fn}\)) simply as an artifact of sampling effort.

\begin{figure}
\hypertarget{fig:geometric}{%
\centering
\includegraphics{./figures/fig1.png}
\caption{\textbf{(A)} The probability that an observed interaction is a
true negative (y-axis) given how many times it has been sampled as a
non-interaction (x-axis). Each color reflects a different value of
\(p_{fn}\), the false-negative-rate (FNR)---this is effectively the cdf
of the geometric distribution. \textbf{(B)} The expected number of total
observations needed (colors) to observe 10 co-occurrences between a
species with relative abundance \(P(A)\) (x-axis) and a second species
with relative abundance \(P(Y)\). \textbf{(C)}: false-negative-rate
(y-axis) as a function of total sampling effort (x-axis) and network
size, computed using the method described above. For 500 independent
draws from the niche model (Williams \& Martinez (2000)) at varying
levels of species richness (colors) with connectance drawn according to
the flexible-links model (MacDonald \emph{et al.} (2020)) as described
in the main text. For each draw from the niche model, 200 sets of 1500
observations are simulated, for which the mean false-negative-rate at
each observation-step is computed. Means denoted with points, with 1 in
the first shade and 2 in the second. \textbf{(D)}: Same as \textbf{(C)},
except using empirical food webs from Mangal database, where richness.
The outlier on \textbf{(D)} is a 714 species
food-web.}\label{fig:geometric}
}
\end{figure}

@@

where the maximum \emph{a posteriori} (MAP) estimate of \((\mu, \phi)\)
applied to Mangal data from (MacDonald \emph{et al.} 2020) is
\((\mu=0.086, \phi=24.3)\). All simulations were done with 500
independent replicates of unique niche model networks per unique number
of observations \(n\). All analyses presented here are done in Julia
v1.8 (Bezanson \emph{et al.} 2015) using both EcologicalNetworks.jl v0.5
and Mangal.jl v0.4 (Banville \emph{et al.} 2021) and are hosted on
\href{https://github.com/gottacatchenall/ms_false_negatives/tree/main/src}{Github}).
Note that the empirical data, for the reasons described above, very
likely already contains many false-negatives, we'll revisit this issue
in the final section.

From fig.~\ref{fig:geometric}(C) it is evident that the number of
species considered in a study is inseparable from the
false-negative-rate in that study, and this effect should be taken into
account when designing samples of ecological networks in the future. We
see a similar qualitative pattern in fig.~\ref{fig:geometric}(D) where
the FNR drops off quickly as a function of observation effort, mediated
by total richness. The practical consequence of the bottom row of
fig.~\ref{fig:geometric} is whether the total number of observations of
all species (the x-axis) for the threshold FNR we deem acceptable (the
y-axis) is feasible. This raises two points: first, empirical data on
interactions are subject to the practical limitations of funding and
human-work hours, and therefore existing data tend to fall on the order
of hundreds or thousands observations of individuals per site. Clear
aggregation of data on sampling effort has proven difficult to find and
a meta-analysis of network data and sampling effort seems both pertinent
and necessary, in addition to the effects of aggregation of interactions
across taxonomic scales (Gauzens \emph{et al.} 2013; Giacomuzzo \&
Jordán 2021). This inherent limitation on in-situ sampling means we
should optimize where we sample across space so that for a given number
of samples, we obtain the maximum information possible. Second, what is
meant by ``acceptable'' FNR? This raises the question: does a shifting
FNR lead to rapid transitions in our ability inference and predictions
about the structure and dynamics of networks, or does it produce a
roughly linear decay in model efficacy? We explore this in the next
section.

@@

\begin{figure}
\hypertarget{fig:resampling_concept}{%
\centering
\includegraphics{./figures/uncertainty_sampler.png}
\caption{(A) The process for estimating the false-negative-rate (FNR)
for an interaction dataset consisting of \(N\) total observed
interactions. (B) The method for resampling interaction probability
based on estimates of false-negative and false-positive rates. (C) The
method for interaction probability resampling applied to three mammals
and three parasites from the Hadfield \emph{et al.} (2014)
dataset.}\label{fig:resampling_concept}
}
\end{figure}

@@

\hypertarget{positive-associations-in-co-occurrence-increase-the-false-negative-rate}{%
\section{Positive associations in co-occurrence increase the
false-negative-rate}\label{positive-associations-in-co-occurrence-increase-the-false-negative-rate}}

The model above doesn't consider the possibility that there are positive
or negative associations which shift the probability of species
cooccurrence together due to their interaction (Cazelles \emph{et al.}
2016). However, here we demonstrate that the probability of having a
false-negative can be higher if there is some positive association in
the occurrence of species \(A\) and \(B\). If we denote the probability
that we observe the co-occurrence of two species \(A\) and \(B\) as
\(P(AB)\) and if there is no association between the marginal
probabilities of observing \(A\) and observing \(B\), denoted \(P(A)\)
and \(P(B)\) respectively, then the probability of observing their
co-occurrence is the product of the marginal probabilities for each
species, \(P(AB) = P(A)P(B)\). In the other case where there is some
positive strength of association between observing both \(A\) and \(B\)
because this interaction is ``important'' for each species, then the
probability of observation both \(A\) and \(B\), \(P(AB)\), is greater
than \(P(A)P(B)\) as \(P(A)\) and \(P(B)\) are not independent and
instead are positively correlated, i.e.~\(P(AB)> P(A)P(B)\). In this
case, the probability of observing a single false-negative in our naive
model from fig.~\ref{fig:geometric}(A) is \(p_{fn}= 1-P(AB)\), which due
to the above inequality implies \(p_{fn}>1-P(A)P(B)\). This indicates an
increasingly greater probability of a false negative as the strength of
association gets stronger, \(P(AB) \to P(AB) \gg P(A)P(B)\). However,
this still does not consider variation in species abundance in space and
time (Poisot \emph{et al.} 2015). If positive or negative associations
between species structure variation in the distribution of \(P(AB)\)
across space/time, then the spatial/temporal biases induced by data
collection would further impact the realized false-negative-rate, as the
probability of false negative would not be constant for each pair of
species across sites.

To test for these positive associations in data we scoured Mangal for
datasets with many spatial or temporal replicates of the same system,
which led the the resulting seven datasets set in figure
fig.~\ref{fig:mangal}. For each dataset, we compute the marginal
probability \(P(A)\) of occurrence of each species \(A\) across all
networks in the dataset. For each pair of interacting species \(A\) and
\(B\), we then compute and compare the probability of co-occurrence if
each species occurs independently, \(P(A)P(B)\), to the empirical joint
probability of co-occurrence, \(P(AB)\). Following our analysis above,
if \(P(AB)\) is greater than \(P(A)P(B)\), then we expect our neutral
estimates of the FNR above to underestimate the realized FNR. In
fig.~\ref{fig:mangal}, we see the difference between \(P(AB)\) and
\(P(A)P(B)\) for the seven suitable datasets with enough spatio-temporal
replicates and a shared taxonomic backbone (meaning all individual
networks use common species identifiers) found on Mangal to perform this
analysis. Further details about each dataset are reported in
tbl.~\ref{tbl:id}.

@@

Here, we assess the effect of false-negatives on our ability to make
predictions about interactions, as well as their effect on network
structure. The prevalence of false-negatives in data is the catalyst for
interaction prediction in the first place, and as a result methods have
been proposed to counteract this bias (Stock \emph{et al.} 2017; Poisot
\emph{et al.} 2022). However, it is feasible that the FNR in a given
dataset is so high that it could induce too much noise for an
interaction prediction model to detect the signal of possible
interaction between species.

@@

\begin{figure}
\hypertarget{fig:addedfnr}{%
\centering
\includegraphics{./figures/fig3.png}
\caption{\textbf{(A)} The area-under the receiver-operator curve
(ROC-AUC) and \textbf{(B)} The area-under the precision-recall curve
(PR-AUC; right) for each different predictive model (colors/shapes)
across a spectrum of the proportion of added false-negatives (x-axis).
\textbf{(C)} The mean trophic-level of all species in a network
generated with the niche model across different species richnesses
(colors). For each value of the FNR, the mean trophic level was computed
across 50 replicates. The shaded region for each line is one
standard-deviation across those replicates.}\label{fig:addedfnr}
}
\end{figure}
